---
title: "Web mapping --- the easy way"
author: "P. Roudier"
date: "26 June 2017"
output: 
  html_document:
    toc: true
    toc_float: true 
    theme: yeti
---

First, let's load our friends: `sf` (vector data), `raster` (raster data), and `dplyr` (general data tidying):

```{r libs, message=FALSE, warning=FALSE, results='hide'}
library(sf)
library(raster)
library(dplyr)
```

Then let's load the `mapview` package, which makes the production of web maps in R very easy:

```{r libs2, message=FALSE, warning=FALSE, results='hide'}
library(mapview)
```

# Datasets

For this demonstration, we are using the `cookfarm` dataset from the `GSIF` package. This dataset is a list with 6 somponents storing various measurements that have been done on the farm.

```{r data}
data('cookfarm', package = "GSIF")

names(cookfarm)
```

## Vector data

The `profiles` component stores soil profile data as a `data.frame`. We will turn this into a simple feature using the `sf` package:

```{r format_data_profiles, message=FALSE, warning=FALSE}
profiles <- cookfarm$profiles %>% 
  st_as_sf(
    coords = c('Easting', 'Northing'), 
    crs = cookfarm$proj4string
  )

# simple plot using the sf package
plot(profiles)
```

## Raster data

The `grids` component is storing gridded data. We will reansform these into a `RasterStack` using the `rasterFromXYZ` function. This function requires the `x` and `y` columns to be the first column of the `data.frame` (from the left) -- a good opportunity to use the `select` function from the `dplyr` package:

```{r format_data_grids, message=FALSE, warning=FALSE}
grids <- cookfarm$grids %>% 
  select(x, y, DEM, TWI, Cook_fall_ECa, Cook_spr_ECa) %>%  # Re-order the columns, and select 4 interesting variables
  rasterFromXYZ(crs = cookfarm$proj4string) # The coordinate reference system is stored as one of the component in cookfarm

# simple plot using the raster package
plot(grids)
```

# Interactive maps using `mapview`

## Vector data

The `mapview` package is basically a wrapper around the `leaflet` package. It makes is very easy to create a `leaflet` map. You can basically just use it as you would use `plot`:

```{r mapview_simple, warning=FALSE}
# couldn't be easier to create a map!
mapview(profiles)
```

But while the `mapview` function can be called just as is, there's numerous details that can be tweaked. For example the colours:

```{r mapview_palette, warning=FALSE}
# Create palettes 
library(RColorBrewer)
pal_continuous <- colorRampPalette(brewer.pal(7, "BrBG")) # For continuous data
pal_categorical <- colorRampPalette(brewer.pal(9, "Set1")) # For categorical data

# Pass the palette to mapview
mapview(
  profiles, 
  zcol = "TAXSUSDA", 
  col.regions = pal_categorical, 
  legend = TRUE
)
```

In particular, a wide range of background (web) maps is available. You can pick one or several. You can find a list of these background from [the Leaflet-extras project website](http://leaflet-extras.github.io/leaflet-providers/preview/): `http://leaflet-extras.github.io/leaflet-providers/preview/`.

```{r mapview_back}
# Tweaking backgrounds
mapview(profiles, zcol = "BLD", col.regions = pal_continuous, legend = TRUE, map.types = "Esri.WorldImagery")
```

The `burst` option can be interesting when you visualise categorical data:

```{r mapview_burst}
# Burst to separate soil classes
mapview(profiles, zcol = "TAXSUSDA", col.regions = pal_categorical, legend = TRUE, burst = TRUE)
```

## Raster data

The `mapview` option also work for data loaded using the `raster` package. If you try to visualise a `RasterStack` (as opposed to `RasterLayer`), you can select and choose which layer to plot using the selection interface.

```{r mapview_grids}
# Plot a RasterLayer
mapview(grids$TWI, col.regions = pal_continuous, na.color = "transparent", legend = TRUE)

# Plot a RasterStack
mapview(grids)

# The same sort of options are available
mapview(grids, col.regions = pal_continuous, na.color = "transparent",  legend = TRUE)
```

### Extent of rasters

You can visualise the raster extent --- rather than the data itself --- using `viewExtent`:

```{r extent}
viewExtent(grids)
```

## Imagery data

For imagery, it is a little diffferent, because usually one wants to visualise a RGB composite of 3 layers, rather than the individual layers individually. In this case, the `viewRGB` is the way to go:

```{r mapview_sat}
# Imagery specific functions
viewRGB(poppendorf)
# The combination of bands can be changed very easily
viewRGB(poppendorf, 4, 3, 2)
```

## Adding `mapviews` together

Conveniently, maps can be aded to each other using the `+` operator:

```{r mapview_both}
# Create 2 maps
m1 <- mapview(grids$DEM, col.regions = pal_continuous, legend = TRUE) 
m2 <- mapview(profiles, zcol = "TAXSUSDA", col.regions = pal_categorical)

# Plot both together
m1 + m2
```

## Visualising projected data

What `mapview` does behind the scenes is to change the projection system to EPSG:3857 (web mercator). In some cases, this is inconvenient and you might actually want to visualise your data in a local projected CRS. `plainview` is here to help:

```{r plainview, eval=FALSE}
plainview(grids$DEM)
```

# Advanced topics

## What happens behind the scenes

`mapview` is a wrapper around `leaflet` package, which is a R API for the popular Javascript library for web mapping called `Leaflet`. `Leaflet` has been designed with simplicity and rapidity in mind. For more power, you'll have to learn a bit more about the `leaflet` package itself. Their website is a great starting point.

## Changing default options

A bunch of options can be changed using the `mapviewOptions` function: 

```{r options}
mapviewOptions(
  basemaps = c("Esri.WorldImagery", "Thunderforest.Landscape"),
  na.color = "transparent"
)

mapview(profiles, zcol = "BLD") + mapview(grids$DEM)
```

## Sync'ing maps

You can associate and synchronise a set of maps using the `sync` function:

```{r mapview_sync, eval=FALSE}
# Syncing several maps
m1 <- mapview(grids$DEM) 
m2 <- mapview(grids$TWI) 
m3 <- mapview(grids$Cook_fall_ECa)

sync(m1, m2, m3, ncol = 2, sync.cursor = TRUE)
```

This is an interactive analogue to the panelled graphs provided by `ggplot2` or `lattice`.

## Slideview

Another advanced visualisation tool is `slideview`, which is convenient to compare two maps:

```{r slideview, eval=FALSE}
img1 <- poppendorf[[1]]
img2 <- poppendorf[[5]]

slideview(
  img1, 
  img2,
  label1 = "Poppendorf-Layer-1",
  label2 = "Poppendorf-Layer-2",
  legend = TRUE
)
```

## Popups

The popups can be either a table (`popupTable`, default behaviour), an image (`popupImage`), or a `htmlwidget` (`popupGraph`).

```{r advanced_mapview_1}
# Table
mapview(
  profiles, 
  popup = popupTable(profiles, zcol = 1:2)
)

# Image
mapview(
  profiles, 
  popup = popupImage('https://www.vcard.wur.nl/WebServices/GetMedia.ashx?id=37263')
)
```

## Garnish map

The `leaflet` and `leaflet.extras` are providing a **LOT** of different map widgets. The `garnishMaps` function facilitates their integration with `mapview`:

```{r garnish}
library(leaflet)

m <- mapview(profiles)
garnishMap(
  m,
  addMouseCoordinates,
  addGraticule,
  addScaleBar
)
```

## Save map

```{r save, eval=FALSE}
# Create map
m  <- mapview(profiles)

# Save interactive HTML page
mapshot(m, url = 'my_map.html')

# Save static image (PNG, JPEG, or PDF)
mapshot(m, file = 'my_image.png')
```

****

# Super advanced `mapview` `r set.seed(4);emo::ji('magic')`

```{r advanced_mapview_2, eval=FALSE}
library(xts)
library(dygraphs)

profiles$SOURCEID <- as.character(profiles$SOURCEID)

records <- cookfarm$readings
records$SOURCEID <- as.character(records$SOURCEID)

ids <- unique(records$SOURCEID)

# Subset sensors
ids <- sample(ids, size = 5) 
  
idx_sensors <- which(profiles$SOURCEID %in% ids)
sensors <- profiles[idx_sensors,]

make_ts <- function(id) {
  
  records %>% 
    filter(SOURCEID == id) %>% 
    dplyr::select(-SOURCEID) %>%
    dplyr::select(Date, ends_with('VW')) %>% 
    xts(.$Date)
}

make_dygraph <- function(id){
  ts <- make_ts(id)
  dygraph(ts)
} 

l_graphs <- lapply(
  ids,
  make_dygraph
)
```

```{r dygraph, eval=FALSE}
make_dygraph(ids[1])
```

```{r result_mapview_dygraph, eval=FALSE}
mapview(sensors, popup = popupGraph(graphs = l_graphs, width = 300, height = 300))
```